Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2017-10-24 aggregation/agg-JIE.tex
morning tweaks

\title{Aggregated Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy}
\author{Brandon Kuczenski\\
Institute for Social, Behavioral, and Economic Research\\
University of California, Santa Barbara}
\date{\mydate\today}

@@

Objectives in critical review vary with the study and the reviewer, but they generally include consistency with the ISO 14044 standard, the validity of methods, appropriateness of data used, correctness of interpretations, and transparency of reporting \citep[Section 6.1]{iso14044}.  Although the object of critical review is typically a written report, the use of a structured disclosure of study design introduces the possibility that reviewers can inspect the model itself and even verify the computations and reported results.  This section discusses how a disclosure in the format described above can enable robust review of a product system model.  Note that, for purposes of critical review, the objective for data privacy can be omitted because the reviewer may be granted privileged access to confidential data.  We return to the data privacy requirement in the next section.

>>>>

\title{Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy}
\author{Brandon Kuczenski\\
Institute for Social, Behavioral, and Economic Research\\
University of California, Santa Barbara}
\date{\mydate\today}

@@

Objectives in critical review vary with the study and the reviewer, but they generally include consistency with the ISO 14044 standard, the validity of methods, appropriateness of data used, correctness of interpretations, and transparency of reporting \citep[Section 6.1]{iso14044}.  Although the object of critical review is typically a written report, the use of a structured disclosure of a PSM introduces the possibility that reviewers can inspect the model itself and even verify the computations and reported results.  This section discusses how a disclosure in the format described above can enable robust review of a product system model.  Note that, for purposes of critical review, the objective for data privacy can be omitted because the reviewer may be granted privileged access to confidential data.  We return to the data privacy requirement in the next section.

### 2017-10-24 aggregation/agg-lca-study.tex
morning tweaks

For clarity, I will call the augmented $\tilde{A}$ and $\tilde{B}$, in which the foreground content always appears to the left of the background content; the first column delivers the study's functional unit; and the background does not depend on the foreground; to be a \emph{canonical LCA foreground study}, for which $\tilde{\mathbf{y}}$ is the canonical functional unit.  This formulation can be applied to the vast majority of LCA case studes, and very likely to all case studies prepared using commercial LCA software.  However, it does not apply for so-called ``integrated hybrid'' models, in which all quadrants contain non-zero cells (for instance, \cite{Gibon_2016}).

It is currently common practice for background database maintainers to pre-compute the life cycle inventory results for their databases.  Using the notation $B_x = B \times (I - A)^{-1}$ to represent the pre-computed background LCI database, Eq.s~\ref{eqn:foreground} can be rewritten as:
\begin{equation}
 \tilde{A} = \left[\begin{array}{cc} 
A_f &  0 \\
A_d & 0 
   \end{array}\right];\;\;\;  \tilde{B} = \left[\begin{array}{cc} B_f & B_x \end{array}\right ]
\label{eqn:fg_bg}
\end{equation}
The proof of the equivalency of Eq.s~\ref{eqn:foreground} and \ref{eqn:fg_bg} is provided in the supplementary materials.  In this form, the main content of the LCIDB has been removed from the $\tilde{A}$ matrix, and the only information remaining there pertains to the study foreground.  Furthermore, the final demand vector has been replaced with a canonical functional unit and no longer needs to be explicitly reported.

>>>>

For clarity, I will call the augmented $\tilde{A}$ and $\tilde{B}$, in which the foreground content always appears to the left of the background content; the first column delivers the study's functional unit; and the background does not depend on the foreground; to be an \emph{LCA foreground study}, for which $\tilde{\mathbf{y}}$ is the canonical functional unit.  This formulation can be applied to the vast majority of LCA case studes, and very likely to all case studies prepared using commercial LCA software.  However, it does not apply for so-called ``integrated hybrid'' models, in which all quadrants contain non-zero cells (for instance, \cite{Gibon_2016}).

It is currently common practice for background database maintainers to pre-compute the life cycle inventory results for their databases.  Using the notation $B_x = B \times (I - A)^{-1}$ to represent the pre-computed background LCI database, Eq.s~\ref{eqn:foreground} can be rewritten as:
\begin{equation}
 \tilde{A}_{flat} = \left[\begin{array}{cc} 
A_f &  0 \\
A_d & 0 
   \end{array}\right];\;\;\;  \tilde{B}_{flat} = \left[\begin{array}{cc} B_f & B_x \end{array}\right ]
\label{eqn:fg_bg}
\end{equation}
The proof of the equivalency of Eq.s~\ref{eqn:foreground} and \ref{eqn:fg_bg} is provided in the supplementary materials.  In this form, the main content of the LCIDB has been removed from the $\tilde{A}$ matrix, and the only information remaining in $\tilde{A}_{flat}$ pertains to the study foreground.  Furthermore, the final demand vector has been replaced with a canonical functional unit and no longer needs to be explicitly reported.

### 2017-10-24 aggregation/proofs.tex
morning tweaks

We begin by computing $(I-\tilde{A})^{-1}$, again taking advantage of the matrix's block triangularity.  Find $Q$ such that:
\begin{equation}
  \left[\begin{array}{cc} (I-A_f) & 0 \\ -A_d & I\end{array}\right]\times
  \left[\begin{array}{cc} (I-A_f)^{-1} & 0 \\ Q & I\end{array}\right] =
  \left[\begin{array}{cc} I & 0 \\ 0 & I\end{array}\right]
\end{equation}

@@

Substituting this into the LCA system equation:
\begin{align}
  s =\;& \mathbf{e}^T\times \tilde{B}\times(I-\tilde{A})^{-1}\tilde{\mathbf{y}}\\
  =\;& \mathbf{e}^T\times \left[\begin{array}{cc} B_f & B_x \end{array}\right]\times
  \left[\begin{array}{cc} (I-A_f)^{-1} & 0 \\ A_d\times(I-A_f)^{-1} & I \end{array}\right]
  \times\left[\begin{array}{c} \tilde{\mathbf{y}}_f \\ \mathbf{0}\end{array}\right]
\end{align}
where $\tilde{\mathbf{y}}_f$ is a canonical functional unit having the same dimension as $A_f$.  Thus:
\begin{align}
s =\;& \mathbf{e}^T\times \left[\begin{array}{cc} B_f & B_x \end{array}\right]\times
 \left[\begin{array}{c} (I-A_f)^{-1}\times\tilde{\mathbf{y}}_f \\ A_d\times(I-A_f)^{-1}\times\tilde{\mathbf{y}}_f\end{array}\right]\\
 =\;& \mathbf{e}^T\times \left(B_f\times(I-A_f)^{-1}\times\tilde{\mathbf{y}}_f
 + B_x\times A_d\times(I-A_f)^{-1}\times\tilde{\mathbf{y}}_f\right)\\
 =\;& \mathbf{e}^T\times \left(B_f + B_x\times A_d\right)\times(I-A_f)^{-1}\times \tilde{\mathbf{y}}_f\\
 =\;& \mathbf{e}^T\times \left(B_f + B_x\times A_d\right)\times \tilde{\mathbf{x}}
\end{align}
thus completing the derivation.

>>>>

We begin by computing $(I-\tilde{A}_{flat})^{-1}$, again taking advantage of the matrix's block triangularity.  Find $Q$ such that:
\begin{equation}
  \left[\begin{array}{cc} (I-A_f) & 0 \\ -A_d & I\end{array}\right]\times
  \left[\begin{array}{cc} (I-A_f)^{-1} & 0 \\ Q & I\end{array}\right] =
  \left[\begin{array}{cc} I & 0 \\ 0 & I\end{array}\right]
\end{equation}

@@

Substituting this into the LCA system equation:
\begin{align}
  s =\;& \mathbf{e}^T\times \tilde{B}_{flat}\times(I-\tilde{A}_{flat})^{-1}\tilde{\mathbf{y}}\\
  =\;& \mathbf{e}^T\times \left[\begin{array}{cc} B_f & B_x \end{array}\right]\times
  \left[\begin{array}{cc} (I-A_f)^{-1} & 0 \\ A_d\times(I-A_f)^{-1} & I \end{array}\right]
  \times\left[\begin{array}{c} \tilde{\mathbf{y}}_f \\ \mathbf{0}\end{array}\right]
\end{align}
where $\tilde{\mathbf{y}}_f$ is a canonical functional unit having the same dimension as $A_f$.  Thus:
\begin{align}
s =\;& \mathbf{e}^T\times \left[\begin{array}{cc} B_f & B_x \end{array}\right]\times
 \left[\begin{array}{c} (I-A_f)^{-1}\times\tilde{\mathbf{y}}_f \\ A_d\times(I-A_f)^{-1}\times\tilde{\mathbf{y}}_f\end{array}\right]\\
 =\;& \mathbf{e}^T\times \left(B_f\times(I-A_f)^{-1}\times\tilde{\mathbf{y}}_f
 + B_x\times A_d\times(I-A_f)^{-1}\times\tilde{\mathbf{y}}_f\right)\\
 =\;& \mathbf{e}^T\times \left(B_f + B_x\times A_d\right)\times(I-A_f)^{-1}\times \tilde{\mathbf{y}}_f\\
 =\;& \mathbf{e}^T\times \left(B_f + B_x\times A_d\right)\times \tilde{\mathbf{x}}
\end{align}
thus completing the derivation.
